Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-29 TACAS2018/paper.tex
ch 6

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs
\footnote{This restricts the observability of the state-space $\vec{x}$, since we can only measure $\vec{y}$}:
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t).
\end{equation}
%
%In this work, 
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals, whose 
duration we refer to as the sample time (cf. Appendix~\ref{sec:continuous_LTI}), 
into the difference equation 
\begin{align}
\label{eq:plant}
\left\{ 
\begin{array}{l}
\vec{x}_{k+1} = \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} = \mat{C}_d \vec{x}_k + \mat{D}_d \vec{u}_k, 
\end{array}
\right.
\end{align} 
%
where $k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$ is the initial state. 
Matrices $\mat{A}_d$, $\mat{B}_d$, $\mat{C}_d$ and $\mat{D}_d$ are equal-sized to $\mat{A}$, $\mat{B}$, $\mat{C}$ and $\mat{D}$ respectively,  
and charactirize the discretized plant dynamics. 
%, whereas the latter denote the continuous evolution.  
In this work we raise the standard assumption that $\mat{D}_d=\mat{D}=\mat{0}$ \cite{Astrom08}.  
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

@@

 \begin{theorem}
 The solution to the differential equation $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$,
 evaluated at time $T$ is
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x} (0)\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(2)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn2}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&\quad gm(\lambda_i)>1\\0&\quad gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  
%$\vec{x}_T$ is the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x} \text{ with } T \in \mathbb{R}^+_0$ is the state of the system at time $t=T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in studying models with parametric inputs, 
 which are described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where $\forall t\geq 0, \vec{u}(t)=\vec{u}$. 
 For such models, a state $\vec{x}_T$ is computed next.  
\begin{theorem}
The state of the model $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$ is 
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u} \text{ where } \mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B} 
 \end{align}
 where $\forall t \leq T,\ \vec{u}(t)=\vec{u}$, and $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}
 \footnote{when $\mat{A}$ is not invertible, we use $\left[\begin{array}{cc}\mat{A}_T&\mat{B}_T\\ 0&\mat{I}\end{array}\right]=e^{\left[\begin{array}{cc}\mat{A}&\mat{B}\\0&0\end{array}\right]}$}.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and given a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, 
 then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}_0+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}_d\vec{u}) \text{ with } T \in \mathbb{R}_0^+$ (where $\mat{A}_d^{\frac{T}{T_d}} = e^{log(\mat{A}_d) \frac{T}{T_d}} = e^{\mat{A} T}$) is the state of the system at time $T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}

>>>>

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs
\footnote{This restricts the observability of the state-space $\vec{x}$, since we can only measure $\vec{y}$. 
\note{what does this mean? do you assume observability? don't you need that if you later compute the observable canonical form?}}:
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t).
\end{equation}
%
\note{what is the dimension of C, D?}
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals, whose 
duration we refer to as the sample time (cf. Appendix~\ref{sec:continuous_LTI}), 
into the difference equation 
\begin{align}
\label{eq:plant}
\left\{ 
\begin{array}{l}
\vec{x}_{k+1} = \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} = \mat{C}_d \vec{x}_k + \mat{D}_d \vec{u}_k, 
\end{array}
\right.
\end{align} 
%
where $k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$ is the initial state. 
Matrices $\mat{A}_d$, $\mat{B}_d$, $\mat{C}_d$ and $\mat{D}_d$ are equal-sized to $\mat{A}$, $\mat{B}$, $\mat{C}$ and $\mat{D}$ respectively,  
and charactirize the discretized plant dynamics. 
%, whereas the latter denote the continuous evolution.  
In this work we raise the standard assumption that $\mat{D}_d=\mat{D}=\mat{0}$ \cite{Astrom08}.  
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

@@

 \begin{theorem}
 The solution to the differential equation $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$,
 \note{matrices S are not defined}
 evaluated at time $T$ is
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x} (0)\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(2)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn2}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&\quad gm(\lambda_i)>1\\0&\quad gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.  
%$\vec{x}_T$ is the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T. 
\note{so: the statement of this theorem is a standard result in the control literature - why is this a new theorem? we do not prove anything new here.}
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x} \text{ with } T \in \mathbb{R}^+_0$ is the state of the system at time $t=T$. \note{there needs to be an x0}
 \note{so I assume this is the new contribution?}
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in studying models with parametric inputs, 
 which are described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where $\forall t\geq 0, \vec{u}(t)=\vec{u}$. 
 For such models, a state $\vec{x}_T$ is computed next.  
\begin{theorem}
The state of the model $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$ is 
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u} \text{ where } \mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B} 
 \end{align}
 where $\forall t \leq T,\ \vec{u}(t)=\vec{u}$, and $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}
 \footnote{when $\mat{A}$ is not invertible, we use $\left[\begin{array}{cc}\mat{A}_T&\mat{B}_T\\ 0&\mat{I}\end{array}\right]=e^{\left[\begin{array}{cc}\mat{A}&\mat{B}\\0&0\end{array}\right]}$}.
 \end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and given a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, 
 then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}_0+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}_d\vec{u}) \text{ with } T \in \mathbb{R}_0^+$ (where $\mat{A}_d^{\frac{T}{T_d}} = e^{log(\mat{A}_d) \frac{T}{T_d}} = e^{\mat{A} T}$) is the state of the system at time $T$.
 \end{corollary}
\begin{proof}
See Appendix~\ref{sec:proofs}
\end{proof}
